First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-08-12 referee_response.txt

Dear Editor and Referee,

We would like to thank the referee for providing a helpful review of our manuscript “First Season MWA EoR Power Spectrum Results at Redshift 7” by myself and others (Submission AAS01756). Below we provide responses to all comments from the referee. The most significant changes are:

- A new figure showing the 2D noise levels to correspond to the 2D power plots we show (Fig. 2).
- A correction to the residual power improvement when using a diffuse model (discussed below).
- Updated references.
- Many expanded explanations on items pointed out by the referee.

We feel the changes have strengthened the article and made it more accessible to the intended audience.

On behalf of the coauthors,

Adam Beardsley
Arizona State University

======================================================================
Specific responses
======================================================================

### 2016-08-22 ms.tex
Removed bold faced changes in prep for arxiv submission.

The 21 cm hyperfine transition from neutral hydrogen residing in the IGM during reionization 
offers another promising method to study the EoR. Not only is the 21cm signal a 
\emph{direct} probe of the IGM, but, due to the narrow width of the transition and the 
relationship between observed frequency and line-of-sight distance, it can be used to map 
the full three dimensional space of the epoch. (For reviews, see \citealt{Furlanetto:2006}, 
\citealt{Morales:2010}, \citealt{Zaroubi:2013}, \citealt{Loeb:2013}, and \citealt{Pritchard:2012}) The first generation of instruments with primary science goals 
to detect the highly redshift 21cm signal have been built, including the GMRT (Giant Metrewave 
Radio Telescope, \citealt{Paciga:2013}), LOFAR (LOw Frequency Array\footnote{http://
www.lofar.org}, \citealt{vanHaarlem:2013, Yatawatta:2013}), PAPER (\textbf{Donald C. Backer} Precision Array for Probing the Epoch of 
Reionization\footnote{http://eor.berkeley.edu}, \citealt{Parsons:2010}), and the MWA 
(Murchison Widefield Array\footnote{http://www.mwatelescope.org}, \citealt{Tingay:2013, Bowman:2013}). 
Due to relatively low signal to noise, these instruments aim for a statistical measurement of 
the EoR in the form of a cosmological power spectrum. 
Complementary to power spectrum experiments are efforts to detect the sky-average global
21-cm signal from the EoR \citep[e.g.][]{Bowman:2010, Sokolowski:2015, Patra:2013, Voytek:2014}.
Meanwhile, the second generation of
21cm EoR interferometers is on its way with the Hydrogen Epoch of Reionization Array (HERA
\footnote{http://reionization.org}, \citealt{DeBoer:2016}), 
and the low frequency Square Kilometer Array (SKA1-Low \citealt{Mellema:2013}),
which will further 
refine the power spectrum measurement in early build-out stages, but will ultimately be 
capable of imaging the ionized bubbles of reionization in its later stages 
\citep{Beardsley:2015,Malloy:2013}.

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}.
\textbf{Furthermore,} \citet{Pober:2015} \textbf{and \citet{Greig:2016} were} able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}, 
while others are investigating the spectral structure of point sources themselves \cite[e.g.][]{Offringa:2016}.
The EoR window is still expected to be preserved above the wedge, and 
several analysis pipelines are under active development 
\textbf{to exploit this foreground isolation}
\citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2015b, Dillon:2013, Trott:2014}.

@@

The data for this work includes two sets. First we use a set of 94 two-minute high band observations of 
the EoR0 field from August 23, 2013, which were taken early in telescope operations and 
found to be particularly well behaved. The MWA EoR collaboration has designated this set 
as a ``golden data set'', which is used to build analysis tools and compare early results 
\textbf{\cite[e.g.][]{Jacobs:2016, Trott:2016, Pober:2016, Dillon:2015}.} All 
techniques demonstrated here will use this golden data set. In Section~\ref{sec:deep} we 
will use a larger data set to move toward a deep integration. This data set consists of all 
EoR0 high band observations beginning August 23, 2013 (including the golden data set), 
and concluding when the field was no longer accessible for the season: when the sun was 
above the horizon for most or all of the field's transit. The final observation of this set was 
on November 29, 2013. The individual snapshots are 112 seconds of data at 0.5 second, 
40 kHz resolution. In total we analyze 2,780 snapshots, or about 86.5 hours of data.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
80~kHz frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citet{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[\textbf{B.~J.}~Hazelton et al. 2016, in prep.]{Barry:2016}.

@@

Under these assumptions we can express the measured, uncalibrated visibility for an 
antenna pair $(i,j)$ as
\begin{equation}
V'_{ij}(\nu) \approx g_i(\nu)g^*_j(\nu)V_{ij}(\nu),
\end{equation}
where $g_i(\nu)$ and $g_j(\nu)$ are the complex gains for antennas $i$ and $j$ 
respectively,
\textbf{and $V_{ij}(\nu)$ is the true visibility which we aim to recover through calibration.}
Because we are treating polarizations independently, we allow the antenna 
subscript to run over both East-West (E-W) and North-South (N-S) polarizations.

@@

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the \textbf{``holographic frame,"} which has been shown to be the optimal \textbf{weighting scheme} for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citet{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz. 
\textbf{This is done to reduce the data volume, and has little impact on our results because
our sensitivity to the EoR is extremely low at the corresponding line-of-sight $k$ modes
\citep{Beardsley:2013}}

@@

\textbf{The pipeline as described up to this point is applied to both the calibrated data
and the foreground model. At this point we subtract the model from the data to form
residual data as well. All three sets of data (dirty, model, and residual) are carried through 
the remaining steps to form corresponding power spectra.}

After squaring and dividing by our observation window function \citep{Bowman:2006}, we 
arrive in three dimensional power spectrum space. We use the even/odd interleaved cubes 
to form a signal (odd plus even) and noise (odd minus even) power spectrum, which we 
subtract from one another to form an unbiased estimate of our signal power. This is 
mathematically equivalent to cross multiplying the even and odd cubes. The three 
dimensional power spectrum cube can next be averaged in annuli 
\textbf{of constant $k_{\perp}$ and $k_{||}$ (i.e. orthogonal to the $k_{||}$ axis)}
to form two dimensional 
power spectra, or spherical shells to form one dimensional power spectra where we will 
ultimately constrain the EoR.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{example_ps.pdf}
\caption{
An example two dimensional, E-W polarization residual power spectrum formed using the 
three hour golden data set
from August 23, 2013. The foreground wedge dominates the region below the solid 
black horizon line, while regions above are mostly noise-like with the exception of horizontal 
coarse band harmonic lines. 
The vertical streaks at high $k_{||}$ and high $k_{\perp}$ are due to sparse $(u,v)$ sampling
beyond the dense core of the MWA.
\textbf{The right axis shows the delay (Fourier dual to frequency) corresponding to the 
$k_||$ axis, and the top axis shows the baseline length (in wavelengths) corresponding
to the $k_{\perp}$ axis.}
\label{fig:example_ps}
}
\end{center}
\end{figure}

An example two dimensional residual power spectrum formed from the three hour golden 
data set is shown in Figure~\ref{fig:example_ps}. The bulk of the residual power is in the 
so-called foreground ``wedge", indicated with the diagonal black lines where the solid line 
corresponds to sky emission at the horizon, and the dashed lines corresponds to emission 
at the edge of the MWA field of view. Above the wedge we see horizontal lines of 
contamination due to the periodic frequency sampling function. These lines are often 
referred to as the coarse band harmonics. We see vertical streaks at high $k_{||}$ and high 
$k_{\perp}$. These are due to sparse $(u,v)$ sampling beyond the dense core of the MWA 
(starting $\sim70\lambda$), 
\textbf{which results in non-uniform spectral sampling after gridding. This in turn
causes foreground power to mix to high spectral modes} \citep{Bowman:2009}. 
\textbf{The frequency dependent sampling creates a point spread function in the $k_{||}$
for each $(u,v)$ cell. In principle covariance weighting along the frequency dimension
may be able to mitigate this leakage \citep[e.g.][]{Liu:2011}, and an \eppsilon implementation 
is under development.}
\textbf{We provide axes more closely related to the measurements to assist in connecting
these instrumental effects. The right axis shows delay (Fourier dual to frequency), and the
top axis shows baseline length measured in wavelengths.}
Between the coarse band lines and to the left of the vertical streaks we see regions 
which contain both positive and negative (indicated by blue on the color bar) pixels. These 
regions are where we have successfully isolated the foregrounds and retained a noise-like 
EoR window in the three hour integration.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{example_2derror.pdf}
\caption{
\textbf{
Uncertainty level corresponding to Figure~\ref{fig:example_ps}. The noise is relatively low at
short baseline length (low $k_{\perp}$) due to the dense MWA core. The array is more
sparse at longer baselines, and the noise increases. The second $k_{||}$ bin has slightly
lower noise than the rest of the $k_{||}$ modes due to the interaction between the
Lomb-Scargle periodogram with the Blackman-Harris window function applied to the data.
}
\label{fig:example_2derror}
}
\end{center}
\end{figure}

\textbf{Figure~\ref{fig:example_2derror} shows the uncertainty level corresponding to the two 
dimensional power spectrum in Figure~\ref{fig:example_ps}. These levels
are found by propagating noise measured from calibrated visibilities through the entire
pipeline. The dense core of the MWA results in relatively low noise at low $k_\perp$.
As the $(u,v)$ density of the array decreases at longer baselines, the noise increases.
The noise is mostly constant across $k_{||}$ because the thermal noise amplitude is
relatively constant over frequency and the Fourier transform distributes the noise power
across all modes \citep{Morales:2004}.
The exception is the second $k_{||}$ bin, which has slightly lower noise than the rest of the 
$k_{||}$ modes. 
As explained in B.~J.~Hazelton et al. (in prep.) this is a result of the Lomb-Scargle periodogram 
interacting with the Blackman-Harris window function we apply to the data. Most of the bins 
have the noise equally distributed between cosine and sine terms, with the exception of the 
$k_{||}=0$ bin where the cosine term carries all the information. Due to the correlations 
introduced by the Blackman-Harris window function the first non-zero $k_{||}$ mode is highly 
covariant with the bottom bin -- therefore the noise on this mode's cosine term is lower than 
other higher order cosines.
}

@@

The top panel of Figure~\ref{fig:cables} shows a 2D power spectrum using the gain model 
described so far. Three of the horizontal lines in the EoR window can be attributed to the 
coarse band gaps, as they reside at the harmonics expected for 1.28~MHz periodic 
sampling. But the sampling cannot account for the fourth line, highlighted by the arrow. The 
corresponding delay time of the line, $\tau \approx 1.23$~$\mu$s, corresponds almost 
exactly to twice the signal travel time through the 150 meter cables (with velocity 0.81c). We 
therefore introduced a reflection term into our gain model, which allows our restricted gains 
to account for the interference of the incident signal with a round-trip reflected signal in the 
cable. 
A similar approach was taken at lower frequencies in \citet{Ewall-Wice:2016}.
The full gain model is expressed as
\begin{equation}\label{eq:cal2}
\hat{g}_i(\nu)=B_{\alpha}(\nu)\left(P_i(\nu)+R_i(\nu)\right),
\end{equation}
where the reflection term can be further decomposed as
\begin{equation}
R_i(\nu) = R_{i,0} e^{-2\pi i \tau_{i} \nu}.
\end{equation}
The reflection coefficient, $R_{i,0}$ is allowed to be complex, while the reflection delay, 
$\tau_i$ is real. After fitting for all other parameters, we fit for the reflection mode. 
\textbf{We found that if we did not first fit and remove the polynomial terms, our reflection
fits would be dominated by the large power in the smooth structure, resulting in solutions that
did not match physical reality.
The reflection fitting is}
done for antennas with suspected reflections, the most offensive of which is seen in the 
antennas with cables of 150 meters. Because the cables were not cut at exact lengths 
(variations on order tens of centimeters), we solve for both the reflection coefficient and 
delay by performing a direct Fourier transform to a highly over resolved delay grid and 
selecting the mode where the reflection amplitude is largest. 
This fitting produces a single reflection coefficient for each antenna with nominal cable
length of 150 meters.

@@

\textbf{This effect is distinct from more fundamental wedge effects 
\citep[e.g.][]{Morales:2012, Hazelton:2013} in that it is a result of the computational limitation
of the analysis. The primary contribution to the foreground mode mixing is due to
the loss of information as each baseline samples a range of $(u,v)$ modes. On the other
hand, the effect discussed here is due to small positional offsets in the gridding kernel
and only exists after gridding.}

@@

Our pipeline uses two modes of FHD -- the full deconvolution mode to identify point sources 
and build a catalog, and a production mode where we calibrate using the sky model and 
subtract it from the data without further fitting. Because full deconvolution on every 
observation from the MWA is computationally not feasible, we currently restrict ourselves to 
the golden data set to build our model. The process of building the catalog is presented in 
\citet{Carroll:2016}. They applied machine learning classification methods 
to select reliable detections from the full deconvolution FHD mode, culminating in the KGS 
catalog\footnote{\textbf{KGS is an abbreviation of KATALOGSS, the KDD (Knowledge 
Discovery in Databases) Astrometry, Trueness, and Apparent Luminosity of Galaxies in 
Snapshot Surveys}}. The work here uses an early iteration of the KGS catalog that was available at the 
time of analysis.

It has been shown that subtracting a foreground model strictly within the main lobe of the 
primary beam will not be sufficient to suppress the power spectrum wedge and unlock the 
EoR window \citep{Thyagarajan:2015b, Thyagarajan:2015, Pober:2016}. 
We are in the process of repeating the model building described above using additional
MWA observations pointed away from our field of interest.
Until that work is complete we supplement the extent of 
our point source model using additional catalogs. We accomplish this through a hierarchical 
catalog pulling from our early KGS catalog, the MWA Commissioning Survey\footnote{The 
more complete GLEAM Survey \citep[][\textbf{N.}~Hurley-Walker et al., in review]
{Wayth:2015} was not available at the time of analysis.} \citep[MWACS, ][]
{Hurley-Walker:2014}, the Culgoora catalog \citep{Slee:1977}, and the Molonglo Reference Catalog 
\citep[MRC, ][]{Large:1981}. Source flux densities are prioritized in this order based on our 
confidence in their predicted flux density at 182 MHz. We first cluster the source lists to avoid 
redundant sources, using a 3.5 arcmin neighborhood radius, then select the flux density from the 
highest priority catalog. Spectral indices were obtained from the MWACS and Culgoora 
catalogs when available, otherwise a two-point spectral index was estimated for Culgoora-
MRC matches or MRC-SUMSS matches. All other sources were given a spectral index of 
$-0.8$, \textbf{the previously determined median spectral index of sources below 1.4 GHz
\citep{Oort:1988, Hunstead:1991, DeBreuck:2000, Mauch:2003}.}
The spectral index was used to extrapolate the flux density from the catalog frequency to 
182 MHz, but a uniform spectral index of $-0.8$ is used within our band when forming 
model visibilities for calibration and foreground subtraction.

@@

Because we use the input foreground model to calibrate our visibilities, the gain estimates 
for our two runs differed. In particular we saw the overall flux scale of the calibrated 
visibilities was higher when using the full hierarchical catalog,
\textbf{owing to a more complete sky model requiring lower amplitude gains to describe the data. 
This difference is on order a part in $10^3$ in mK$^2$ units, but is largely amplified
when observing the difference in the power spectra.}
In order to put both residual 
2D power spectra onto the same scale for comparison, we first divide each 
\textbf{$(k_{\perp},k_{||})$ cell}
by the 
corresponding dirty power spectrum pixel, then subtract to arrive at a ratio difference. 
\begin{equation}\label{eq:diff_ratio}
\mathtt{ratio\;difference = \frac{residual_1}{dirty_1} - \frac{residual_2}{dirty_2}}
\end{equation}
Here we use subscript 1 to represent the MWACS catalog spectra, and 2 for the 
hierarchical catalog spectra. The ratio difference is shown in Figure~\ref{fig:mwacs_ratio}. 
The entire wedge being positive (blue) demonstrates that using the new catalog resulted in 
successfully subtracting higher fractional power. Not only does this indicate we subtract 
more power, but it also confirms that our calibrated data is more closely matched to the 
model, meaning that our calibration solutions in general are more accurate.

>>>>

The 21 cm hyperfine transition from neutral hydrogen residing in the IGM during reionization 
offers another promising method to study the EoR. Not only is the 21cm signal a 
\emph{direct} probe of the IGM, but, due to the narrow width of the transition and the 
relationship between observed frequency and line-of-sight distance, it can be used to map 
the full three dimensional space of the epoch. (For reviews, see \citealt{Furlanetto:2006}, 
\citealt{Morales:2010}, \citealt{Zaroubi:2013}, \citealt{Loeb:2013}, and \citealt{Pritchard:2012}) The first generation of instruments with primary science goals 
to detect the highly redshift 21cm signal have been built, including the GMRT (Giant Metrewave 
Radio Telescope, \citealt{Paciga:2013}), LOFAR (LOw Frequency Array\footnote{http://
www.lofar.org}, \citealt{vanHaarlem:2013, Yatawatta:2013}), PAPER (Donald C. Backer Precision Array for Probing the Epoch of 
Reionization\footnote{http://eor.berkeley.edu}, \citealt{Parsons:2010}), and the MWA 
(Murchison Widefield Array\footnote{http://www.mwatelescope.org}, \citealt{Tingay:2013, Bowman:2013}). 
Due to relatively low signal to noise, these instruments aim for a statistical measurement of 
the EoR in the form of a cosmological power spectrum. 
Complementary to power spectrum experiments are efforts to detect the sky-average global
21-cm signal from the EoR \citep[e.g.][]{Bowman:2010, Sokolowski:2015, Patra:2013, Voytek:2014}.
Meanwhile, the second generation of
21cm EoR interferometers is on its way with the Hydrogen Epoch of Reionization Array (HERA
\footnote{http://reionization.org}, \citealt{DeBoer:2016}), 
and the low frequency Square Kilometer Array (SKA1-Low \citealt{Mellema:2013}),
which will further 
refine the power spectrum measurement in early build-out stages, but will ultimately be 
capable of imaging the ionized bubbles of reionization in its later stages 
\citep{Beardsley:2015,Malloy:2013}.

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}.
Furthermore, \citet{Pober:2015} and \citet{Greig:2016} were able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}, 
while others are investigating the spectral structure of point sources themselves \cite[e.g.][]{Offringa:2016}.
The EoR window is still expected to be preserved above the wedge, and 
several analysis pipelines are under active development 
to exploit this foreground isolation
\citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2015b, Dillon:2013, Trott:2014}.

@@

The data for this work includes two sets. First we use a set of 94 two-minute high band observations of 
the EoR0 field from August 23, 2013, which were taken early in telescope operations and 
found to be particularly well behaved. The MWA EoR collaboration has designated this set 
as a ``golden data set'', which is used to build analysis tools and compare early results 
\cite[e.g.][]{Jacobs:2016, Trott:2016, Pober:2016, Dillon:2015}. All 
techniques demonstrated here will use this golden data set. In Section~\ref{sec:deep} we 
will use a larger data set to move toward a deep integration. This data set consists of all 
EoR0 high band observations beginning August 23, 2013 (including the golden data set), 
and concluding when the field was no longer accessible for the season: when the sun was 
above the horizon for most or all of the field's transit. The final observation of this set was 
on November 29, 2013. The individual snapshots are 112 seconds of data at 0.5 second, 
40 kHz resolution. In total we analyze 2,780 snapshots, or about 86.5 hours of data.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
80~kHz frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citet{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[B.~J.~Hazelton et al. 2016, in prep.]{Barry:2016}.

@@

Under these assumptions we can express the measured, uncalibrated visibility for an 
antenna pair $(i,j)$ as
\begin{equation}
V'_{ij}(\nu) \approx g_i(\nu)g^*_j(\nu)V_{ij}(\nu),
\end{equation}
where $g_i(\nu)$ and $g_j(\nu)$ are the complex gains for antennas $i$ and $j$ 
respectively,
and $V_{ij}(\nu)$ is the true visibility which we aim to recover through calibration.
Because we are treating polarizations independently, we allow the antenna 
subscript to run over both East-West (E-W) and North-South (N-S) polarizations.

@@

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the ``holographic frame," which has been shown to be the optimal weighting scheme for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citet{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz. 
This is done to reduce the data volume, and has little impact on our results because
our sensitivity to the EoR is extremely low at the corresponding line-of-sight $k$ modes
\citep{Beardsley:2013}.

@@

The pipeline as described up to this point is applied to both the calibrated data
and the foreground model. At this point we subtract the model from the data to form
residual data as well. All three sets of data (dirty, model, and residual) are carried through 
the remaining steps to form corresponding power spectra.

After squaring and dividing by our observation window function \citep{Bowman:2006}, we 
arrive in three dimensional power spectrum space. We use the even/odd interleaved cubes 
to form a signal (odd plus even) and noise (odd minus even) power spectrum, which we 
subtract from one another to form an unbiased estimate of our signal power. This is 
mathematically equivalent to cross multiplying the even and odd cubes. The three 
dimensional power spectrum cube can next be averaged in annuli 
of constant $k_{\perp}$ and $k_{||}$ (i.e. orthogonal to the $k_{||}$ axis)
to form two dimensional 
power spectra, or spherical shells to form one dimensional power spectra where we will 
ultimately constrain the EoR.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{example_ps.pdf}
\caption{
An example two dimensional, E-W polarization residual power spectrum formed using the 
three hour golden data set
from August 23, 2013. The foreground wedge dominates the region below the solid 
black horizon line, while regions above are mostly noise-like with the exception of horizontal 
coarse band harmonic lines. 
The vertical streaks at high $k_{||}$ and high $k_{\perp}$ are due to sparse $(u,v)$ sampling
beyond the dense core of the MWA.
The right axis shows the delay (Fourier dual to frequency) corresponding to the 
$k_||$ axis, and the top axis shows the baseline length (in wavelengths) corresponding
to the $k_{\perp}$ axis.
\label{fig:example_ps}
}
\end{center}
\end{figure}

An example two dimensional residual power spectrum formed from the three hour golden 
data set is shown in Figure~\ref{fig:example_ps}. The bulk of the residual power is in the 
so-called foreground ``wedge", indicated with the diagonal black lines where the solid line 
corresponds to sky emission at the horizon, and the dashed lines corresponds to emission 
at the edge of the MWA field of view. Above the wedge we see horizontal lines of 
contamination due to the periodic frequency sampling function. These lines are often 
referred to as the coarse band harmonics. We see vertical streaks at high $k_{||}$ and high 
$k_{\perp}$. These are due to sparse $(u,v)$ sampling beyond the dense core of the MWA 
(starting $\sim70\lambda$), 
which results in non-uniform spectral sampling after gridding. This in turn
causes foreground power to mix to high spectral modes \citep{Bowman:2009}. 
The frequency dependent sampling creates a point spread function in the $k_{||}$
for each $(u,v)$ cell. In principle covariance weighting along the frequency dimension
may be able to mitigate this leakage \citep[e.g.][]{Liu:2011}, and an \eppsilon implementation 
is under development.
We provide axes more closely related to the measurements to assist in connecting
these instrumental effects. The right axis shows delay (Fourier dual to frequency), and the
top axis shows baseline length measured in wavelengths.
Between the coarse band lines and to the left of the vertical streaks we see regions 
which contain both positive and negative (indicated by blue on the color bar) pixels. These 
regions are where we have successfully isolated the foregrounds and retained a noise-like 
EoR window in the three hour integration.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{example_2derror.pdf}
\caption{
Uncertainty level corresponding to Figure~\ref{fig:example_ps}. The noise is relatively low at
short baseline length (low $k_{\perp}$) due to the dense MWA core. The array is more
sparse at longer baselines, and the noise increases. The second $k_{||}$ bin has slightly
lower noise than the rest of the $k_{||}$ modes due to the interaction between the
Lomb-Scargle periodogram with the Blackman-Harris window function applied to the data.
\label{fig:example_2derror}
}
\end{center}
\end{figure}

Figure~\ref{fig:example_2derror} shows the uncertainty level corresponding to the two 
dimensional power spectrum in Figure~\ref{fig:example_ps}. These levels
are found by propagating noise measured from calibrated visibilities through the entire
pipeline. The dense core of the MWA results in relatively low noise at low $k_\perp$.
As the $(u,v)$ density of the array decreases at longer baselines, the noise increases.
The noise is mostly constant across $k_{||}$ because the thermal noise amplitude is
relatively constant over frequency and the Fourier transform distributes the noise power
across all modes \citep{Morales:2004}.
The exception is the second $k_{||}$ bin, which has slightly lower noise than the rest of the 
$k_{||}$ modes. 
As explained in B.~J.~Hazelton et al. (in prep.) this is a result of the Lomb-Scargle periodogram 
interacting with the Blackman-Harris window function we apply to the data. Most of the bins 
have the noise equally distributed between cosine and sine terms, with the exception of the 
$k_{||}=0$ bin where the cosine term carries all the information. Due to the correlations 
introduced by the Blackman-Harris window function the first non-zero $k_{||}$ mode is highly 
covariant with the bottom bin -- therefore the noise on this mode's cosine term is lower than 
other higher order cosines.

@@

The top panel of Figure~\ref{fig:cables} shows a 2D power spectrum using the gain model 
described so far. Three of the horizontal lines in the EoR window can be attributed to the 
coarse band gaps, as they reside at the harmonics expected for 1.28~MHz periodic 
sampling. But the sampling cannot account for the fourth line, highlighted by the arrow. The 
corresponding delay time of the line, $\tau \approx 1.23$~$\mu$s, corresponds almost 
exactly to twice the signal travel time through the 150 meter cables (with velocity 0.81c). We 
therefore introduced a reflection term into our gain model, which allows our restricted gains 
to account for the interference of the incident signal with a round-trip reflected signal in the 
cable. 
A similar approach was taken at lower frequencies in \citet{Ewall-Wice:2016}.
The full gain model is expressed as
\begin{equation}\label{eq:cal2}
\hat{g}_i(\nu)=B_{\alpha}(\nu)\left(P_i(\nu)+R_i(\nu)\right),
\end{equation}
where the reflection term can be further decomposed as
\begin{equation}
R_i(\nu) = R_{i,0} e^{-2\pi i \tau_{i} \nu}.
\end{equation}
The reflection coefficient, $R_{i,0}$ is allowed to be complex, while the reflection delay, 
$\tau_i$ is real. After fitting for all other parameters, we fit for the reflection mode. 
We found that if we did not first fit and remove the polynomial terms, our reflection
fits would be dominated by the large power in the smooth structure, resulting in solutions that
did not match physical reality.
The reflection fitting is
done for antennas with suspected reflections, the most offensive of which is seen in the 
antennas with cables of 150 meters. Because the cables were not cut at exact lengths 
(variations on order tens of centimeters), we solve for both the reflection coefficient and 
delay by performing a direct Fourier transform to a highly over resolved delay grid and 
selecting the mode where the reflection amplitude is largest. 
This fitting produces a single reflection coefficient for each antenna with nominal cable
length of 150 meters.

@@

This effect is distinct from more fundamental wedge effects 
\citep[e.g.][]{Morales:2012, Hazelton:2013} in that it is a result of the computational limitation
of the analysis. The primary contribution to the foreground mode mixing is due to
the loss of information as each baseline samples a range of $(u,v)$ modes. On the other
hand, the effect discussed here is due to small positional offsets in the gridding kernel
and only exists after gridding.

@@

Our pipeline uses two modes of FHD -- the full deconvolution mode to identify point sources 
and build a catalog, and a production mode where we calibrate using the sky model and 
subtract it from the data without further fitting. Because full deconvolution on every 
observation from the MWA is computationally not feasible, we currently restrict ourselves to 
the golden data set to build our model. The process of building the catalog is presented in 
\citet{Carroll:2016}. They applied machine learning classification methods 
to select reliable detections from the full deconvolution FHD mode, culminating in the KGS 
catalog\footnote{KGS is an abbreviation of KATALOGSS, the KDD (Knowledge 
Discovery in Databases) Astrometry, Trueness, and Apparent Luminosity of Galaxies in 
Snapshot Surveys}. The work here uses an early iteration of the KGS catalog that was available at the 
time of analysis.

It has been shown that subtracting a foreground model strictly within the main lobe of the 
primary beam will not be sufficient to suppress the power spectrum wedge and unlock the 
EoR window \citep{Thyagarajan:2015b, Thyagarajan:2015, Pober:2016}. 
We are in the process of repeating the model building described above using additional
MWA observations pointed away from our field of interest.
Until that work is complete we supplement the extent of 
our point source model using additional catalogs. We accomplish this through a hierarchical 
catalog pulling from our early KGS catalog, the MWA Commissioning Survey\footnote{The 
more complete GLEAM Survey \citep[][N.~Hurley-Walker et al., in review]
{Wayth:2015} was not available at the time of analysis.} \citep[MWACS, ][]
{Hurley-Walker:2014}, the Culgoora catalog \citep{Slee:1977}, and the Molonglo Reference Catalog 
\citep[MRC, ][]{Large:1981}. Source flux densities are prioritized in this order based on our 
confidence in their predicted flux density at 182 MHz. We first cluster the source lists to avoid 
redundant sources, using a 3.5 arcmin neighborhood radius, then select the flux density from the 
highest priority catalog. Spectral indices were obtained from the MWACS and Culgoora 
catalogs when available, otherwise a two-point spectral index was estimated for Culgoora-
MRC matches or MRC-SUMSS matches. All other sources were given a spectral index of 
$-0.8$, the previously determined median spectral index of sources below 1.4 GHz
\citep{Oort:1988, Hunstead:1991, DeBreuck:2000, Mauch:2003}.
The spectral index was used to extrapolate the flux density from the catalog frequency to 
182 MHz, but a uniform spectral index of $-0.8$ is used within our band when forming 
model visibilities for calibration and foreground subtraction.

@@

Because we use the input foreground model to calibrate our visibilities, the gain estimates 
for our two runs differed. In particular we saw the overall flux scale of the calibrated 
visibilities was higher when using the full hierarchical catalog,
owing to a more complete sky model requiring lower amplitude gains to describe the data. 
This difference is on order a part in $10^3$ in mK$^2$ units, but is largely amplified
when observing the difference in the power spectra.
In order to put both residual 
2D power spectra onto the same scale for comparison, we first divide each 
$(k_{\perp},k_{||})$ cell
by the 
corresponding dirty power spectrum pixel, then subtract to arrive at a ratio difference. 
\begin{equation}\label{eq:diff_ratio}
\mathtt{ratio\;difference = \frac{residual_1}{dirty_1} - \frac{residual_2}{dirty_2}}
\end{equation}
Here we use subscript 1 to represent the MWACS catalog spectra, and 2 for the 
hierarchical catalog spectra. The ratio difference is shown in Figure~\ref{fig:mwacs_ratio}. 
The entire wedge being positive (blue) demonstrates that using the new catalog resulted in 
successfully subtracting higher fractional power. Not only does this indicate we subtract 
more power, but it also confirms that our calibrated data is more closely matched to the 
model, meaning that our calibration solutions in general are more accurate.
